---
title: "Check-in 1: R Markdown"
author: "Maxim Dokukin"
date: "2024-01-29"
output: html_document
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

1.  This is a code chunk that opens the `mtcars` dataset, which contains information about on 32 cars from a 1974 issue of the magazine *Motor Trend* and creates a scatter plot with horsepower `hp` on the x-axis and miles-per-gallon `mpg` on the y-axis. Try running this code and Rendering/Knitting your Rmd document. If you have done this correctly, RStudio should produce an html document.

```{r, eval = T, echo = T}
data(mtcars)
plot(mtcars$hp, mtcars$mpg)
```

2.  Using this scatter plot, describe the relationship between `mpg` and `hp`.
Higher hp correlates to lower mpg.

3.  Create your own code chunk below this question containing the line of code `print("Hello MATH167R!")`. Then Render/Knit your Rmd document. If you have done this correctly, RStudio should produce an html document.
```{r}
print("Hello MATH167R!")
```
